	I would like to thank my thesis advisor Stefan Ballmer for trusting in me as a scientist, always providing indispensable advice, and allowing me to follow my interests.
	
	To Peter Saulson: Your love of science truly inspires me to be a better scientist and teacher.
	
	To Duncan Brown:  Tough questions force us to grow and group meetings provided a growth spurt.
	
	To Sheila Dwyer, Jenne Driggers, and Georgia Mansell, thank you for your patience in teaching me the gritty details about the Hanford interferometer, I will never forget the difficult in-chamber work on HAM6 and the long commissioning hours in the control room.
	
	To Daniel Sigg, your dinners are the perfect fugue of flavors, conversations and friends. Thank you for letting me participate in some of the crucial interferometer commissioning tasks.
	
	To Keita Kawabe, thank you for the tough questions but also the encouragement to do the best science possible.
	
	To Rick Savage, the Fellows program is one of the best parts of being at Hanford and I was honored to be a part of it for so long, thank you for the advice about all aspects of life and the fishing trips just when I needed it most.
	
	To Aidan Brooks, thank you for letting me work on/use TCS at Hanford, your guidance and experience is invaluable in times of confusion.
	
	To Daniel Vander-Hyde, thank you for helping at Hanford with your extra hands/brain and being patient with me when things weren't going well but also celebrating with me when things were!
	
	To Dan Brown, thanks for coming across the pond to help when we needed it most and the Inbetweeners jokes at wings night.
	
	To Corey Gray, Gunner and Gomez, the hikes brought peace to my troubled mind and patience to my un-restful soul.
	
	To TJ Shaffer: Thank you for helping when we needed steady hands or Guardian skills.
	
	To all the operators and staff at Hanford: Jeff, Betsy, Chandra, TJ, Jim, Cheryl, Nico, Patrick, Ed, Gerardo, Evan.
	
	To Hang Yu: Thank you for helping me understand the inner workings of the interferometer with only a few lines of math and a picture.
	
	To my forever teachers Melissa Brainerd, Paul Didier, Chrissy Dahms, Carolyn West, and Dan Dorsey:  The accumulated effect of my work is a direct result of your teachings, you inspire kids like me from White Center to chase our dreams.
	
	To my childhood friends Quang, Leah, Thang, Sienna, Khanh, Truc, Kay, Tuan, Wanda, James, Matt and Trung:  You knew me when I was just dreaming of being a scientist, now we're here.
	
	To the Bloomingdale family: Thank you for always making me feel at home even in the coldest of Syracuse winters.
	
	To Jaysin Lord, Laura Nuttall, and Mat Hutchens: Thank you for the laughs and Thursday wings.
	
	To my family Cuong, Nhi, Tony, Theresa, Brittny, Kayla, Chloe: Your support and encouragement mean everything to me.
	
	To my parents: The difficulties of science dwindles in comparison to your journey from Vietnam, this work is dedicated to you.
	
	To my darling Tiffany: Your support provides nourishment for my soul and your love gives a guiding light through the darkest of nights.