\chapter{Introduction}

	Say something profound here.

	\section{Gravitational Waves}\label{gravitational waves}
	In 1915, Albert Einstein published his theory of general relativity \ref{einstein}.
	
	The seminal equation in this theory is:
	
	\begin{equation} \label{einstein}
	G_{\mu \nu} = 8 \pi T_{\mu \nu}
	\end{equation}
	
	Which is a set of 10 coupled second-order differential equations that are nonlinear.  In its complete form, equation \ref{einstein} fully describes the interaction between space-time and mass-energy. 
	
	
	To describe the physics in a highly curved space-time, one would have to fully solve the Einstein field equations numerically. In areas where the curvature is close to flat,  the weak field approximation can be applied and the metric is described as	
	
	\begin{equation} \label{weak}
	g_{\mu \nu}  \approxeq \eta_{\mu \nu} + h_{\mu \nu}
	\end{equation}
	
	where $\eta_{\mu \nu}$ is the metric of flat space time and $|h_{\mu \nu}| \ll 1$ is the perturbation due to a gravitational field.
	
	By plugging in equation \ref{weak} into \ref{einstein} and using empty space we obtain the familiar wave equation
	
	\begin{equation} \label{wave}
	\Big(\nabla^2 - \frac{1}{c^2} \pdv[2]{t} \Big) h_{\mu \nu}  = 0
	\end{equation}

	which has a plane-wave solution of the form $h_{\mu \nu} = A_{\mu \nu} e^{ik_{\nu} x^{\nu}}$. 
	
	Using the gauge constraint $h^{\mu \nu}_{,\nu} = 0$, it follows that $A_{\mu \nu} k^{\mu} = 0$ which means that the gravitational wave amplitude is orthogonal to the propagation vector.
	
	Further imposing transverse-traceless gauge and assuming that the wave is traveling in the $x^3$ direction, it can be shown that the complex amplitude has physical significance expressed in the matrix
	
	\begin{equation} \label{gwamp}
	A_{\mu \nu} = 
	\begin{pmatrix}
			0 &    0   &  0      & 0 
		 \\ 0 & A_{xx} &  A_{yx} & 0
		 \\ 0 & A_{xy} & -A_{xx} & 0
		 \\ 0 &    0   &  0      & 0
	\end{pmatrix}
	\end{equation}

	Oftentimes, the four non-zero components of equation \ref{gwamp} can be categorized into two distinct polarizations called plus and cross such that $h_{+} = A_{xx} = -A_{yy}$ and $h_{\cross} = A_{xy} = A_{yx}$ .

 
	It is natural to attempt to understand the physical interpretation of equation \ref{gwamp} as an affect on the position of a free floating particle. Consider the four-velocity, $U^{\alpha}$, in the transverse traceless gauge where the coordinate itself is attached the particles and incorporates any small wiggles that would shake the coordinates.  Of course, any free particles will follow the geodesic equation
	
	\begin{equation}\label{geodesic}
	\nabla U^{\alpha} = \frac{\text{d}}{\text{d} \tau} U^{\alpha} + \Gamma^{\alpha}_{\mu \nu} U^{\mu} U^{\nu} = 0
	\end{equation}	
	where $\Gamma^{\alpha}_{\mu \nu} = \frac{1}{2} g^{\gamma \alpha}(g_{\gamma \mu, \nu} + g_{\gamma \nu,\mu} - g_{\mu \nu, \gamma} )$ are the famous Christoffel symbols.
	By evaluating the first term of the acceleration in equation \ref{geodesic},
	\begin{equation}\label{accel}
	\bigg(\frac{\text{d}U^{\alpha}}{\text{d}\tau}\bigg)_0 = -\Gamma^{\alpha}_{00} 
	\\ = \frac{1}{2} \eta_{\mu \nu} (h_{\beta 0, 0} + h_{0 \beta, 0} + h_{0 0, \beta} )
	\end{equation}
	However, comparing equation \ref{gwamp} and equation \ref{accel}, it is clear that if the particle is initially at rest, then a moment later it is still at rest! The term "at rest" is actually used liberally here since the coordinate system varies along with the gravitational wave. 
	
	Alternatively, one can ask if a gravitational wave passed by a pair of particles separated by length $L$, what would be the effect on the distance between two points?  The proper distance is defined as

	\begin{equation}\label{propdist}
	\delta l
	= \int{g_{\mu \nu} dx^{\mu} dx^{\nu}} \\
	= \int_{0}^{L}{g_{xx} {d}x}\\
	\approx |g_{xx}(x=0)|^{1/2}\\
	\approx [ 1 + \frac{1}{2} h_{xx}(x=0)] L
	\end{equation} 
	
	which shows us two very important points about the nature of gravitational waves.  Firstly, the effect is very small since the length variation, $h_{xx}$, is a small perturbation on flat space-time.  Secondly, the effect is proportional to the initial separation between the particles. This means a detector which is large will have a better chance to measure these small effects, an important point that drove the design of the Laser Interferometer Gravitational-Wave Interferometer (LIGO).
	
	
	\subsection{Energy from a Gravitational Wave}
	Consider a binary system.
	

	\subsection{Measuring Gravitational Waves with Light}\label{measuringGWs}
	Even with the theoretical formulation of gravitational waves resolved by the 1970s ref[Pirani], detection of GWs by ground-based detectors was still a controversial topic among scientists in the field ref[Collins].  This was due to the incredible accuracy required to measure the strain from even the most dense astrophysical objects ref[Rai].  The earliest attempts at detecting these small signals were famously done by Joseph Weber ref[Weber] using large resonant bars and piezoelectric transducers to extract the energy from gravitational waves at the resonant frequencies of the bars. Picture of bar detectors at LHO.  However, these bars are limited by thermal noise and can only detect GWs in very narrow frequency bands ref[Bars]. 
	
	Interferometers are devices that measure small displacements by using a laser that is split with a partially transmitting mirror (or beamsplitter), which allows 50$\%$ of the light to get reflected and 50$\%$ to be transmitted.  Each of the split beams travel down arms and reflect off of mirrors and return down the arms.  Upon reaching the beamsplitter, the beams recombine and by using the superposition of electromagnetic waves the laser will add linearly at the output port (or antisymmetric port).  The laser beams will gather phase as they propagate down each individual arm, and when recombining, the intensity of the light will be proportional to the phase differences between each beam.  This will correspond to a differential length that is described by
	
	\begin{equation}
	L_{-} = l_{x} - l_{y}
	\end{equation}
	
	As shown in equation \ref{propdist}, the effect of gravitational waves on the proper length between two free falling objects is proportional to the initial separation.  From figure[GWparticles], it is intuitively clear that interferometry would be an ideal technique to detect signals from a gravitational wave.  However, one can explicitly derive a ground-based interferometer's response to a GW from an astrophysical object.
	
	Consider a gravitational wave source arbitrarily located in the sky with respect to an interferometer on Earth. By denoting the interferometer's Cartesian coordinates as $\{\hat{x},\hat{y},\hat{z}\}$ with x and y located along the arms respectively such that the z-axis points directly towards the zenith (i.e. a right-handed system.).  Using the well known Euler angles, a relation from the detector frame to the source frame with coordinates $\{\hat{x}',\hat{y}',\hat{z}'\}$ can be seen in Figure[Euler]. 
	
	If a gravitational wave at the source has emitted GWs with plus and cross polarizations as denoted by equation \ref{gwamp}, then the detector time series can be regarded as ref[Duncan Thesis and AndersonGWs]
	\begin{equation}
	h(t) = F_{+}(\theta,\phi,\psi) \, h_{+}(t) + F_{\times}(\theta,\phi,\psi) \, h_{\cross}(t)
	\end{equation}
	where $F_{+}(\theta,\phi,\psi)$ and $F_{\cross}(\theta,\phi,\psi$ are the antenna pattern functions that project the gravitational wave amplitudes onto the detector frame.
	
	\begin{equation}\label{Fplus}
	F_{+}(\theta,\phi,\psi) = -\frac{1}{2}[1+\text{cos}^2(\theta)] \text{cos}(2\phi) \text{cos}(2\psi) - \text{cos}(\theta) \text{sin}(2\phi) \text{sin}(2\psi)
	\end{equation}
	\begin{equation}\label{Fcross}
	F_{\cross}(\theta,\phi,\psi) = + \frac{1}{2}[1+\text{cos}^2(\theta)] \text{sin}(2\phi) \text{cos}(2\psi) - \text{cos}(\theta) \text{sin}(2\phi) \text{sin}(2\psi)
	\end{equation}
	
	If the gravitational wave is located directly above the interferometer (ie $\theta = 0$) and setting $\psi=0$, then the magnitude of the antenna pattern is equal to unity.  Furthermore, by rotating about the $\phi$ angle such that the detector arms align with the plus polarization, the null geodesic equation (ie the path of a photon) in the interferometer frame becomes 
	
	\begin{equation}
	ds^2 = g_{\mu\nu}dx^{\mu} dx^{\nu} = -dt^2 + [1+h_{+}]  dx^2 + [1-h_{+}]  dy^2 + dz^2 = 0
	\end{equation}
	
	Now if the photon is traveling along the x-arm, this means that $dy^2 = dz^2 = 0$ and the metric equation transforms to
	
	\begin{equation}
	\frac{dt}{dx} = \sqrt{ 1+h_{+} } \approx 1+\frac{1}{2} h_{+} 
	\end{equation}
	
	The amount of time required for the photon to reach the x-end mirror (starting at $t=0$) is equal to
	
	\begin{equation}\label{pathlength}
	t_1 = \int_{0}^{L_{x}} [1+\frac{1}{2}  h_{+}(x) ] \text{d}x
	\end{equation}
	
	where $t_0$ is the start time and $L_x$ is the total length of the x-arm.  Upon returning to the beamsplitter, the photon's total time of flight for the x and y arms are, respectively,
	
	\begin{equation}
	t_2 = 2 L_x + \frac{1}{2} \int_{0}^{L_x} \bigg[  h_{+}(x) +  h_{+}(x + L_x)  \bigg] \text{d}x
	\end{equation}
	\begin{equation}
	t'_{2}= 2 L_y - \frac{1}{2} \int_{0}^{L_y} \bigg[  h_{+}(y) +  h_{+}(y + L_y)  \bigg] \text{d}y
	\end{equation}
	
	If the gravitational wave period is much longer than the time of flight, then $h_{+}$ does not change much during the measurement, which means $h_{+}(\eta_i) \approx h_{+}(\eta_i + L_{\eta_i}) \approx constant$.  By subtracting the flight times of the photons for each arm and setting $L = L_x = L_y$, the difference is proportional to the gravitational wave perturbation multiplied by the sum of arm lengths (with a factor of $c$ to get the units right),
	
	\begin{equation}
	\Delta t = t_2 - t'_{2} = \frac{2L}{c} h_{+}
	\end{equation}
	
	By recasting the expression for time of flight in terms of the phase picked up laser light as it travels through space, the differential phase shift is
	\begin{equation}\label{diffphase}
	\Delta \Phi = \Phi(t_{2}) - \Phi(t'_{2}) = \frac{4 \pi}{\lambda} \, h_{+} \, L
	\end{equation}
	
	The equation above is simple, however, it only works for gravitational wave signals that are not frequency dependent and it assumes that the path length can be arbitrarily long. Both points are actually not true but we can alleviate these discrepancies by considering a gravitational wave signal of the form $h(t) = h_0 \text{exp} (i 2 \pi f_{GW} t)$  and repeating the calculation between equations \ref{pathlength} and \ref{diffphase}:
	
	\begin{equation}\label{gwsinc}
	\Delta \Phi (t) = h(t) \; \tau \; \frac{2 \pi c}{\lambda} \; \text{sinc}(f_{GW} \tau) e^{i \pi f_{GW} \tau}
	\end{equation}
	
	where $\tau_{RT} = 2L/c$.  The response for a detector whose arm lengths are blank km have null points at the around blank Hz which means the instrument cannot be arbitrarily long, however, this point is not concerning because it is too expensive and difficult to make a terrestrial detector of this size.
	
	How to practically measure $\Delta \Phi$ with an interferometer is explained in section \ref{LIGOInstrument}.

	\subsection{Detection of Gravitational Waves}
	The LIGO project is meant to observe astrophysical objects so it is natural to wonder how well a single detector can probe the universe.  
	
	Signal-to-noise is the level of interesting signals compared to the background noise, which can be expressed as
	\begin{equation}\label{SNR}
	\rho = 2 \int_{-\infty}^{\infty} \frac{ \tilde{s}(f) \tilde{s}^*(f) }{S_n(f)}
	\end{equation}
	
	where ${S_n(f)}$ is the one-sided average power spectral density of the detector noise and $\tilde{s}(f)$ is the Fourier transform of the detectors' response to a gravitational-wave signal.
	
	Consider two dense objects with mass $m_1$ and $m_2$ rotating around each other, separated by a distance $R$ such that quadrupole equation still holds (i.e. flat space-time with small perturbations). The detector response, in units of strain, to the gravitational waves emitted by this source is [Finn],
	
	\begin{equation}\label{inspiralsignal}
	s(t) = \frac{\mathcal{M}}{d_L} \Theta(\theta,\phi,\psi,i) [\pi f(t) \mathcal{M}]^{2/3} \cos[\Phi(t) + constant]
	\end{equation}
	
	where $\mathcal{M} = (1+z) \frac{(m_1 m_2)^{3/5}}{(m_1 + m_2)^{1/5}}$ is the chirp mass, $d_L$ is the luminosity distance. 
	
	
	The orientation response, $\Theta(\theta,\phi,\psi,i)$, is a function that depends entirely on the detector orientation relative to the angular momentum vector of the binary [Figure of a binary spiraling and a detector on earth]
	
	\begin{equation}
	\Theta(\theta,\phi,\psi,i) = 2 \sqrt{	[F_{+}(\theta,\phi,\psi) (1+\cos^2(i)) ]^2 + [2 F_{\cross} \cos(i)]^2 }
	\end{equation}
	
	where $F_{+}$ and $F_{\cross}$ are from equation \ref{Fplus} and \ref{Fcross}.
	
	As the binary loses energy to gravitational waves, the orbit will shrink [ref Schutz or something] as a function of time, in turn, the orbital frequency will increase
	
	\begin{equation}
	f(t) = \frac{1}{\pi \mathcal{M}} \bigg[\frac{5}{256} \frac{\mathcal{M}}{T-t}\bigg]^{3/8} 
	\end{equation}
	
	By defining the binary phase as $f(t) = \frac{1}{2\pi} \frac{\partial \mathbf{\Phi} }{\partial t}$, it is easy to recognize that the phase also evolves with time.
	\begin{equation}
	\Phi(t) = 2\pi \int_{T}^{t} f(t) dt = -2 \bigg( \frac{T-t}{5\mathcal{M}}\bigg)^{5/8}
	\end{equation}
	
	What this means is that the gravitational wave signal from the binary will increase in amplitude and frequency as a function of time up until the coalescence time, $T$.


	There is actually a subtle difference between the coalescence time and the moment when adiabatic approximations fail which allows usage of the quadrupole formulation. 

	Plugging equation \ref{inspiralsignal} into \ref{SNR},
	\begin{equation}
	\rho = 8 \Theta(\theta,\phi,\psi,i) \frac{r_0}{d_L} \bigg(\frac{\mathcal{M}}{1.2 \textup{M}_\odot}\bigg)^{5/6} \zeta(f_{max})
	\end{equation}
	
	There are two important functions in the equation above that reflect the detector's performance in sensing gravitational radiation from a binary inspiral, $r_0$ and $\zeta(f_{max})$.
	
	The characteristic distance, $r_0$ is how far the detector can see for a fixed binary's mass distribution,
	
	\begin{equation}\label{char_r0}
	r_0^2 = \frac{5}{192\pi^{4/3}} \bigg(\frac{3\textup{M}_\odot}{20}\bigg)^{5/3}   \int_{0}^{\infty} \frac{1}{S_n(f)} \frac{\text{d} f}{f^{7/3}}
	\end{equation}

	Due to the integrand's dependence on $f^{-7/3}$, lower frequency improvements in the detector's noise spectrum will contribute more to the distance.
	
	\begin{equation}\label{zeta}
	\zeta(f_{max}) = \frac{\int_{0}^{2f_{max}} \frac{1}{S_n(f)} \frac{\text{d} f}{f^{7/3}}}{\int_{0}^{\infty} \frac{1}{S_n(f)} \frac{\text{d} f}{f^{7/3}}}
	\end{equation}
	
	The sensitivity can be different for various mass distributions; $\zeta(f_{max})$ is a normalized function that describes how well the detector's bandwidth overlaps the binary's frequency evolution.  If $f_{max}$ is higher that the minimal sensitivity frequency for the detector, then there is good overlap and  $\zeta(f_{max})\approx 1$.  However, if the masses are sufficiently large, $f_{max}$ will be lower because coalescence occurs before the objects reach a higher frequency regime and this will result in  $\zeta(f_{max})\approx 0$.
	
	The process of two objects coalescing starts at lower frequencies and terminate at some $f_{max}$ that depends on when the objects reach their inner-most stable circular orbit, $f_{ISCO}$.   If $m_1=m_2$, then the maximum frequency is[]
	
	\begin{equation}
	\begin{aligned}
	f_{max}	=&  \frac{f_{ISCO}}{1+z} \\
			=&	\frac{99 \text{Hz}}{1+z} \, \frac{20 \textup{M}_\odot}{M}
	\end{aligned}
	\end{equation}
	where $M=m_1+m_2$.
	
	
	By plugging a few mass distributions and an estimate of the sensitivity [GWINC], it is possible to study the range of a single detector.  For example, a 1.4-1.4$\textup{M}_\odot$ binary that has optimal orientation with respect to the detector has horizon distance equal to...
	
	Horizon distance and range, advanced LIGO projected
	
	[Finn, Duncan Thesis]
	\cite{Saulson}
	
	\section{The LIGO Instrument}\label{LIGOInstrument}
	In the simplest form, the LIGO instrument is an incredibly large Michaelson interferometer.  If we imagine the output as a measure of the differential arm length, it becomes a natural way of detecting gravitational waves. However, to make a practical gravitational-wave observatory, the complexity will have to be extended.  The next sections will explore the various upgrades that were implemented to improve the sensitivity of LIGO.
	
		\subsection{Simple Michaelson}\label{michelson}
		As show in figure[michaelson], the interferometer readout uses a photodetector that measure the total power which depends on how laser light in the arms constructively or destructively interferes.  
		At the end of section \ref{measuringGWs}, it was shown that the differential time of flight between photons traveling down each arm carries gravitational wave information.  The difference in flight times, $\Delta t$, can be converted into how much phase,$\Delta \phi$ is accumulated by the photons as they propagate through space.  But the question remains how an interferometer actually measures $\Delta \phi$:
		
		If the input electric field of the interferometer is $E_0$, the beamsplitter will transmit $E_0 /2$ down the x-arm and reflect $E_0 /2$ down the y-arm.  By setting the beamsplitter to be the origin,  two beams traveling down their respective arms will have gathered a phase $\phi_i$. Then, upon reflecting off the end mirrors and returning to the beamsplitter, each of the electric fields can be described by these equations
		
			\begin{equation}
			\begin{aligned}
				E_{x} 	&=	\frac{i E_0}{2} e^{2i\phi_{x}}	
			\\	E_{y} 	&=	\frac{i E_0}{2} e^{2i\phi_{y}}
			\end{aligned}
			\end{equation}
			
		Since the electromagnetic waves are linear, the resultant sum of waves at the output will be $E_{out} = E_x + E_y$. A photodiode (PD) is placed at the output (or antisymmetric) port to read out the integrated power which is related to the total electric field by
		\begin{equation}
		\begin{aligned}\label{asy_power}
			P \vert_{PD}	&= \int_{Area} I \;				\text{d}A 
		\\					&= \int_{Area} (E_x + E_y)^2 \;	\text{d}A 
		\\					&= P_{in} \; \text{cos}^2(\Delta \phi)
		\end{aligned}
		\end{equation}	
		where $\Delta \phi = \phi_{x} - \phi_{y} = k_x L_x - k_y L_y$ and $P_{in}$ is the input power.. By using equation \ref{diffphase} and the common (or average) arm length $L_{+} = \frac{L_x + L_y}{2}$, the differential phase results in a power shift
		
		\begin{equation}
		P \vert_{PD} \approx P_{max} \; (1-2 \Delta \phi) = P_{in} \; (1-2 k L_{+} h_{+})
		\end{equation}
		
		There is a large DC term that is dependent on the input power and generally, it is very difficult to measure small changes in a large signal. So the next obvious method would be to shift the arms such that the output port is operating on a dark fringe, normally this is called a null-point operation.  However, there are difficulties associated with this method as well.  Consider shifting the phase of equation \ref{asy_power} by $\pi/2$, which would result in
		
		\begin{equation}\label{null}
		P \vert_{null} = P_{in} \; \text{sin}^2 (\Delta \phi) \approx P_{in} \; (k L_{+} h_{+})^2 
		\end{equation}
		However, this results in second-order dependence on the gravitational-wave signal which is already approximated to be very small.
		
		So a good solution to the issue of how to $read$ out a gravitational wave signal from an interferometer is familiar to radio engineers, namely, RF detection methods:
		
		Consider changing the interferometer input by adding an electro-optical modulator (EOM) to sinusoidally modulate the laser frequency and expanding to first order using the Bessel functions,
		
		\begin{equation}\label{modE}
		\begin{aligned}
		E_{in} 	&= E_{0} e^{i(wt + \beta \text{cos} (\Omega t))} \\
				&\approx E_0 e^{iwt} [J_0(\beta) + J_1(\beta) e^{i \Omega t} + J_1(\beta) e^{-i \Omega t}] \\
				&= E_{C,in} + E_{SB+,in} + E_{SB-,in}
		\end{aligned}
		\end{equation}
		where $\Omega$ and $\beta$ are the modulation frequency and depth, respectively.  The 
		
		It is useful to define an optical transfer functions which transforms the interferometer's input electric field to its output,
		
		\begin{equation}\label{opt_tf}
		E_{out} = E_{C} + E_{SB+} + E_{SB-} = 
		\begin{pmatrix}
			t_{C} 	&   
		\\ 	t_{SB+} &
		\\ 	t_{SB-} &
		\end{pmatrix}
		\begin{pmatrix}
		E_{C,in} &    E_{SB+,in}    &  E_{SB-,in}     
		\end{pmatrix}
		\end{equation}
		
		
		The carrier transfer function, $t_{C}$ has already been calculated by equations \ref{asy_power} - \ref{null} and the sideband transfer functions are not much different.
		
		\begin{equation}\label{sb_tf}
		t_{SB\pm} = r_{x,\pm}  e^{i\phi_{\pm,x}} - r_{y,\pm}  e^{i\phi_{\pm,y}}
		\end{equation}
		
		where $\phi_{\pm,i} = (k \pm k_{\Omega}) \, \ell_{i} = (\frac{w+\Omega}{c} ) \ell_{i}$. In the current example, there is no difference in reflectivity between the sidebands and carrier fields, however, this will not be true in general when integrating Fabry-Perot cavities.  Plugging equation \ref{sb_tf} into \ref{opt_tf}, the output electric field becomes 
		
		\begin{equation}
		E_{out} = i e^{iwt} [ J_0(\beta) 	k \ell_{+}  h_{+}  \; + \; J_1(\beta) \sin( k \Delta \ell + k_{\Omega} \Delta \ell) e^{i\Omega t} \; + \; J_1(\beta) \sin( k \Delta \ell + k_{\Omega} \Delta \ell) e^{-i\Omega t} ]
		\end{equation}
		
		By choosing the carrier signal to be on the dark fringe, $k \Delta \ell = \pi/2$ but the sidebands to be slightly off the null (Figure []) and hence leak into the anti-symmetric port, the electric field reduces to
		
		\begin{equation}
		E_{out} = i e^{iwt} [ J_0(\beta) 	k \ell_{+}  h_{+}  \; + \; J_1(\beta) \sin(k_{\Omega} \Delta \ell) e^{i\Omega t} \; + \; J_1(\beta) \sin( k_{\Omega} \Delta \ell) e^{-i\Omega t} ]
		\end{equation}
		
		Recall that the intensity is equal to the electric field squared,
		
		\begin{equation}
		\begin{aligned}
			I	= \vert E_{out} \vert^2  =	&\vert E_{C}\vert^2 + \vert E_{SB+}\vert^2 + \vert E_{SB-}\vert^2 \\
										  	& + 2 \mathbf{Re} \{ \; E_{SB+} E^*_{SB-} e^{2i\Omega t} \; \}\\
										  	& + 2 \mathbf{Re} \{ \; (E_{C} E^*_{SB-} +  E_{SB+} E^*_{C} ) e^{i\Omega t} \; \}
		\end{aligned}
		\end{equation}
		
		The last term is referred to as the $beat$ $note$ between the carrier signal and the sidebands.  It is possible to extract only the signal at the modulation frequency using a mixer and low-pass filter to get a demodulated signal.  The mixer is an analog device which takes in two signals and outputs the product of the two inputs. Using a local oscillator at the modulation frequency, the demodulated intensity is 
		
		\begin{equation}
		\begin{aligned}
		I_{Demod} 	&\propto \big[ 4 \pi  J_0(\beta) J_1(\beta) \frac{\ell}{\lambda}  \sin(k_{\Omega} \Delta \ell)  \; h_{+}\big] \big[ \cos(\Omega t)  \sin(\Omega t + \phi_{Demod}) \big] \\
					&= \big[ 4 \pi  J_0(\beta) J_1(\beta) \frac{\ell}{\lambda}  \sin(k_{\Omega} \Delta \ell)  \; h_{+}\big] \big[ \sin(\phi_{Demod}) + \sin(2\Omega t + \phi_{Demod}) \big]
		\end{aligned}
		\end{equation}
		
		where $\phi_{Demod}$ is the phase that can be set by the user in order to practically extract the right signal. For the output, the   After the mixer, there will be signals at DC, $\Omega$, $2\Omega$ and so on. However, the part that is linear in the gravitational wave amplitude will be at DC so a low-pass filter will allow the final signal to dominate:
		
		\begin{equation}
		S = 4 \pi  J_0(\beta) J_1(\beta) \frac{\ell}{\lambda}  \sin(k_{\Omega} \Delta \ell) \sin(\phi_{Demod}) \; h_{+}
		\end{equation}
		
		This shows that a RF detection technique will be linear in GW signal with no large DC offset.  However, this method depends on the local oscillator's stability. Reference [Fritzchel, Hild] 
		
		Setting the carrier on a null point means $\Delta \ell = \frac{k_{\Omega}}{k} \frac{\pi}{2}$ and allows the designer to optimize the Schnupp asymmetry length to get the best signal for some modulation frequency.
		
		This type of readout scheme is called Heterodyne detection, where the local oscillator signal is provided by an EOM.  However, the Advanced LIGO scheme uses Homodyne detection in order to derive the GW signal but 
		
		Disadvantages of heterodyne scheme: fluctuations in the local oscillator via the modulation depth changing or the phase oscillation.
		
		Fix: Homodyne scheme, ref[Tobin and Nick Smith], this has the advantage of being mode-matched to the carrier beam and lower quantum noise and easier to do squeezing. see section []
		
		
		\cite{BlackPDH}	
	
		\subsection{Fabry-Perot Cavities}\label{FP}
		There are two ways to make our instrument better: one is to increase the sensitivity to gravitational waves and the other is to make our noise lower. From equation \ref{diffphase}, the gravitational wave signal is proportional to the optical path length that the photon travels, which means the most straightforward method of increasing the sensitivity is to make the arms as long as possible (up to the null point described by equation \ref{gwsinc}.  Generally, there were two methods to do this: a delay line or a Fabry-Perot resonator.  At the time of writing this thesis, all modern gravitational wave detectors use the latter method.
	
		A Fabry-Perot cavity is an optical system comprised of two mirrors and one laser input. The very simple condition that once the laser has made a round trip around the optics, it is the same shape and size.  Conceptually, this may seem simple but in practice, controlling and sensing any optical cavity comes with many challenges.
		
		Plane wave analysis:
		Consider a two mirror system separated by a length $L$ with reflection and transmission coefficients: $r_1$, $t_1$, $r_2$, $t_2$.
		
		Starting with a plane wave at the input mirror with amplitude $E_0$ \cite{Saulson}, 
		
		the reflection coefficient is 
		\begin{equation}\label{r_FP}
		r_{FP} = -r_1 + \frac{t_1^2 r_2  e^{-i2kL}}{1-r_1 r_2 e^{-i2kL}}
		\end{equation}
		
		the transmission coefficient is
		\begin{equation}\label{t_FP}
		t_{FP} = \frac{t_1 t_2 e^{ikl}}{1-r_1 r_2 e^{-i2kl}}
		\end{equation}
		
		the circulating coefficient is
		\begin{equation}\label{c_FP}
		c_{FP} = \frac{t_1}{1- r_1 r_2 e^{-2ikL}}
		\end{equation}
		
		Resonance Condition: $L = n \lambda / 2$  
		
		If you are on resonance, the circulating coefficient in the cavity is maximized so that the gain is
		
		\begin{equation}
		\text{Gain} = c^2_{FP} \vert_{\text{resonating}} = \bigg( \frac{t_1}{1-r_1 r_2}\bigg)^2
		\end{equation}
		
		Frequency response of a single FP (plot):
		\cite{Lawrence:99} \cite{dynamicFP}
		A cavity of fixed length and frequency, the circulating power becomes
	
		which is maximized when 
		\begin{equation}
		\frac{\Delta w}{w} = -\frac{\Delta L}{L}
		\end{equation}
		
		Notice that the circulating power depends on the length of the cavity as well as the frequency and one might naively determine that modulating the two parameters independently cause the same effect.  However, when both the frequency and the length are changing, the resonant light  they are related by a frequency dependent transfer function
		\begin{equation}
		C(s) \frac{\Delta w}{w} = -\frac{\Delta L}{L}
		\end{equation}
		where
		\begin{equation}
		C(s) = \frac{1-e^{-2sL/c}}{2sL/c}
		\end{equation}
		in the Laplace domain.
		
		Finesse, or the line width of the resonant peak, function of $r_1$ and $r_2$      :
		\begin{equation}
		\mathbb{F} = \frac{\pi \sqrt{r_1 r_2}}{1- r_1 r_2}
		\end{equation}
		
		Storage Time :
		\begin{equation}
		\tau_{s} = \frac{L}{c \pi} \, \mathbb{F}
		\end{equation}
		
		Cavity Pole:
		\begin{equation}
		f_{pole} = \frac{1}{4\pi \tau_{s}}
		\end{equation}
	
		Free Spectral Range:
		\begin{equation}
		f_{FSR}  = \frac{c}{2L}
		\end{equation}
		
		Stability: In order to prove that the Fabry Perot is stable, it is useful to introduce the matrices that describe a periodic optical system which is explicitly derived in appendix \ref{FPappendix}.  A Fabry Perot cavity that is separated by distance $L$ with spherical mirrors that have radii of curvature $R_1$ and $R_2$ will need to satisfy 
		\begin{equation}\label{gfactor}
		0 \geq \bigg(1+\frac{L}{R_1}\bigg) \bigg(1+\frac{L}{R_2}\bigg) \geq 1
		\end{equation}
		in order to be geometrically stable.
		
		Power circulating as a function of our defined parameters slightly off resonance by a length of $\delta L$:
		\begin{equation}
		P_{cav} = \vert c_{FP} \vert^2 = \frac{Gain}{ 1 + \big(\frac{2\mathbb{F}}{\pi} \big)^2 \, \text{sin}^2(k \delta L) }
		\end{equation}
		
		Modal Contents
		
		\subsubsection{Locking a Fabry Perot Cavity}
		This is described everywhere, reference Black, Drever, Jenne, Evan Hall. Just give the highlights.
		
		Described above are the theoretical constructs of a FP cavity, but the question remains, how does one practically construct a resonant optical cavity?  The answer comes from using a heterodyne sensing scheme similar to the one described in section [].  Except the optical system is not a Michelson interferometer but rather a two mirror cavity, however, this can apply to a number of different geometries such as triangular or bow-tie cavities shown in Figure [].  All of which are used in LIGO for various reasons.
		
		Starting with a input laser and an EOM (electro-optical modulator) that imparts upper and lower sidebands at a modulation frequency $\Omega$, the user injects three beams into the optical system exactly as in equation \ref{modE}.  When placing a photodetector on the cavity reflection port, one should see the effects of the cavity on each of the three electric fields.
		
		\begin{equation}
		E_{FP,out} = E_{C} + E_{SB+} + E_{SB-} = 
		\begin{pmatrix}
		r_{C} 	&   
		\\ 	r_{SB+} &
		\\ 	r_{SB-} &
		\end{pmatrix}
		\begin{pmatrix}
		E_{C,in} &    E_{SB+,in}    &  E_{SB-,in}     
		\end{pmatrix}
		\end{equation}
		
		where the reflection coefficients follow equation \ref{r_FP}.  It is obvious that the phase accumulated by the cavity differ between the carrier and sidebands so that the beat frequency will be proportional to the phase difference.  To make a good reference for the resonant carrier beam, the sidebands should be anti-resonant in the cavity.  
		\begin{equation}
		r_{C} = -r_1 + \frac{t_1^2 r_2  e^{-i2kL}}{1-r_1 r_2 e^{-i2kL}}
		\end{equation}
		
		and 
		\begin{equation}
		r_{SB\pm} = -r_1 + \frac{t_1^2 r_2  e^{-i2(k+k_{\Omega})L}}{1-r_1 r_2 e^{-i2(k+k_{\Omega})L}}
		\end{equation}
		
		This time, 
		
		\subsubsection{Application to LIGO}
		
		Frequency response of a FP Michaelson:
		Just write it down.
		References:[Black, Rick's Paper]

		\subsection{Power-Recycled Fabry-Perot Interferometers}
		Power Recycling
		If the interferometer is operating such that the 4 km arms are exactly different in arms a pi over two times the wavelength, then the intensity of the light at antisymmetric port will be close to null.  This means the power from the arms will reflect back towards the input laser.  Ref[] shows the effect of adding a partially reflecting mirror to increase the optical gain of the Michaelson for the sidebands and carrier fields.
		
		
		
		References: [Meers, Kiwamu]
		
		\subsection{Dual-Recycled Fabry-Perot Interferometers}
		
		One of the biggest changes made between initial and advanced LIGO was the addition of a signal recycling mirror at the antisymmetric port shown in Figure []. This extra mirror allows an extra degree of freedom in shaping the sensitivity curve such.
		
		
		Figure comparing the three cases of increased sensitivity.
		
		References: [Kiwamu]
		
		\subsection{Fundamental Noise Sources}
		Ref: Evan Hall, GWINC

		Noise budget:
		
		\subsubsection{Quantum Noise}
		One fundamental noise source that is limiting LIGO's sensitivity arises from the fluctuations of quantum vacuum entering the anti-symmetric port and coupling to the electric fields.  A quantum mechanical description of an interferometer was constructed by Caves[] where he used electric field operators to show that quantum vacuum fluctuations are the cause of radiation pressure and shot noise in an interferometer.
		
		\textbf{Quantum states}:
		
		\begin{equation}
		\hat{N} = \hat{a}^{\dagger} \, \hat{a}
		\end{equation}
		
		
		Unitary operator:
		\begin{equation}
		\hat{D} \equiv \hat{D}(\alpha) \equiv \text{exp}(\alpha \hat{a}^{\dagger} - \alpha^{*} \hat{a} ) = e^{-\frac{\vert{\alpha}\vert^2}{2}} e^{\alpha \hat{a}^{\dagger} } e^{\alpha^{*} \hat{a} }
		\end{equation}

		$\hat{D}(\alpha) = \hat{D}^{-1}(\alpha) = \hat{D}(-\alpha)$
		
		\begin{equation}
		\begin{aligned}
		\hat{D}^\dagger&\, \hat{a} 		\,\hat{D}			= \hat{a} + \alpha \\ 
		\hat{D}^\dagger&\, \hat{a}^\dagger \,\hat{D} 		= \hat{a}^\dagger + \alpha^*
		\end{aligned}
		\end{equation}
		
	
		\begin{equation}
		\ket{\alpha} = \hat{D} \ket{0} =  e^{-\frac{\vert{\alpha}\vert^2}{2}} e^{\alpha \hat{a}^{\dagger} } \ket{0}
		\end{equation}
		
		
		\textbf{Radiation Pressure}
		
		One might naively think that power fluctuations in the laser cause radiation pressure effects on the test masses which will result in noise.  However, if the 50/50 beamsplitter is perfect, then the momentum transfer to each test mass will be a common length changes and will not change the intensity at the antisymmetric port (or the symmetric port for that matter).
		
		Following the work of Caves[Caves], the concept of radiation pressure arises from first considering the plane wave waves from both the symmetric and anti-symmetric ports.  This method is similar to the input-output methods of section \ref{michelson} where the beamsplitter will couple the electric fields from the input laser and quantum vacuum.
		
		The electric fields hitting the test masses, respectively,
		\begin{subequations}\label{exey}
		\begin{equation}
		E_x = \frac{1}{\sqrt{2}} \bigg[ iE_0 +   E_{AS,in} \bigg]
		\end{equation}
		\begin{equation}
		E_y = \frac{1}{\sqrt{2}} \bigg[  E_0 + i E_{AS,in} \bigg]
		\end{equation}
		\end{subequations}
		
		The intensities hitting the test masses, respectively,
		\begin{subequations}
		\begin{equation}
		\vert E_x \vert^2 = \frac{1}{2} \bigg[ \vert E_0 \vert^2 + \vert E_{AS,in} \vert^2  + i(E_0 E^*_{AS,in} - E_0^*E_{AS,in} )\bigg]
		\end{equation}
		\begin{equation}
		\vert E_y \vert^2 = \frac{1}{2} \bigg[ \vert E_0 \vert^2 + \vert E_{AS,in} \vert^2  - i(E_0 E^*_{AS,in} - E_0^*E_{AS,in} )\bigg]`
		\end{equation}
		\end{subequations}
		
		Momentum transfer between the two test masses:
		\begin{equation}\label{momentumtranfer}
		\begin{aligned}
		 \mathbf{P} 	&= \frac{2 \hbar \omega}{c} \bigg( \vert E_x \vert^2 - \vert E_y \vert^2 \bigg) \\
						&= \frac{2 \hbar \omega}{c} \bigg( E_0 E^*_{AS,in} - E_0^*E_{AS,in} \bigg)\\
		\Rightarrow	\mathbf{\hat{P}}&= \frac{2 \hbar \omega}{c} \bigg( \hat{a}_1^{\dagger} \hat{a}_2 - \hat{a}_2^{\dagger} \hat{a}_1 \bigg)
		\end{aligned}
		\end{equation}
		
		The last part of equation \ref{momentumtranfer} replaces the classical electric fields with creation and annihilation operators for the symmetric input mode ($\hat{a}_1^{\dagger}$, $\hat{a}_1$) and antisymmetric input mode ($\hat{a}_2^{\dagger}$, $\hat{a}_2$) modes. Recall that there is a well established convention [Gerry] denoting the wave function of the two dimensional modes of the quantum harmonic oscillator, $\ket{\alpha,\beta}$, where $\alpha$ denotes the input symmetric mode and $\beta$ refers to the input antisymmetric mode.
		
		\begin{equation}
		\ket{\alpha,0} = \hat{D}_1(\alpha) \ket{0,0}
		\end{equation}
		
		
		
		The interesting results arising from this formulation are the expectation value:

		\begin{equation}
		\langle \mathbf{\hat{P}} \rangle = \bra{\alpha,0} \mathbf{\hat{P}}\ket{\alpha,0} = 0
		\end{equation}
		
		And the variance 
		\begin{equation} 
		\begin{aligned}
		\Delta \mathbf{P}^2 &= \langle \mathbf{\hat{P}}^2 \rangle  - \langle \mathbf{\hat{P}} \rangle^2 \\
							&= \bigg(\frac{2 \hbar \omega}{c}\bigg)^2 \bra{\alpha,0}  \hat{a}_1^{\dagger}  \hat{a}_2  \hat{a}_1  \hat{a}_2^{\dagger} + \hat{a}_2^{\dagger}  \hat{a}_1  \hat{a}_2  \hat{a}_1^{\dagger}
							- \hat{a}_2^{\dagger}  \hat{a}_1  \hat{a}_1  \hat{a}_2^{\dagger} - \hat{a}_1^{\dagger}  \hat{a}_2  \hat{a}_2  \hat{a}_1^{\dagger} \ket{\alpha,0}\\
							&= \bigg(\frac{2 \hbar \omega}{c}\bigg)^2 \vert \alpha \vert^2
		\end{aligned}
		\end{equation}
		
		
		For some time interval, input laser power consisting of $\langle N \rangle =\vert\alpha \vert^2$ photons for a time interval $\Delta T$ is,
		\begin{equation}
		P_{in} = \frac{\hbar \omega}{\Delta T} \vert \alpha \vert^2
		\end{equation}
	
		To solve for the amplitude spectral density of the displacement, Newton's second law can be applied in the frequency domain
		\begin{equation}
		M (2\pi f)^2 \tilde{x}(f) = \tilde{F}(f) = \frac{\Delta \mathbf{P}}{\Delta T}
		\end{equation}
		
		The force spectrum is white because the impacting photons arrive randomly.
		
		Solving for the displacement,
		\begin{equation}
		\tilde{x}_{RP}(f) = \sqrt{\frac{\hbar \omega}{\Delta T} P_{in}} \frac{1}{2Mc (\pi f)^2}
		\end{equation}
	
		The noise spectral density for $M=40kg$, $P_{in}=125$W, and $\lambda=1064$nm
		\begin{equation}
		\tilde{h}_{RP}(f) = 2.04\times 10^{-20} \frac{1}{f^2} \; \bigg[ \frac{\text{Strain}}{\sqrt{\text{Hz}}}\bigg]
		\end{equation}
		
		
		%\bf{Shot Noise}
		Another way that quantum fluctuations can vary the antisymmetric port output is by adding phase noise.  Imagine holding the test masses rigidly such that the only effects on the output light is due to a phase change in the laser light in the arms. By using the same formulation for radiation pressure but propagating the fields back to the beam splitter, equations \ref{exey} will have extra phase,
		
		\begin{subequations}\label{exeyout}
		\begin{equation}
		E_{x,out} = \frac{1}{\sqrt{2}} \bigg[ iE_0 +   E_{AS,in} \bigg] e^{-i2kL_x}
		\end{equation}
		\begin{equation}
		E_{y,out} = \frac{1}{\sqrt{2}} \bigg[  E_0 + i E_{AS,in} \bigg] e^{-i2kL_y}
		\end{equation}
		\end{subequations}
		
		Antisymmetric port output electric field is
		\begin{equation}
		\begin{aligned}
		E_{AS,out} 	&= \frac{1}{\sqrt{2}} (E_{x,out} + iE_{y,out})\\
					&= i e^{-ikL_x-ikL_y} \big[\cos(\Delta \phi) E_0 - \sin(\Delta \phi) E_{AS,in}\big]
		\end{aligned}
		\end{equation}
		where $\Delta \phi = k(L_x-Ly)$
		
		Output intensity,
		\begin{equation}
		\begin{aligned}
		\mathbf{I}_{AS,out} 		&= \vert E_{AS,out}\vert^2 \\
									&= \bigg[ \cos^2(\Delta \phi)\vert E_0\vert^2 + \sin^2(\Delta\phi)\vert E_{AS,in}\vert^2 - \sin(\Delta\phi)\cos(\Delta\phi) \big[E_0 E^*_{AS,in} + E_0^* E_{AS,in}\big] \bigg]\\
		\Rightarrow	
		\hat{\mathbf{I}}_{AS,out} 	&= \bigg[ \cos^2(\Delta \phi)a_1^{\dagger}a_1 + \sin^2(\Delta\phi)a_2^{\dagger}a_2 - \sin(\Delta\phi)\cos(\Delta \phi) \big[a_1^{\dagger}a_2 + a_2^{\dagger}a_1 \big] \bigg]\\	
		\end{aligned}
		\end{equation}
		
		The expectation value for the intensity using an input coherent laser with $\alpha$ and quantum vacuum input at the antisymmetric port is
		
		\begin{equation}
		\begin{aligned}
		\langle \hat{\mathbf{I}} \rangle 	&= \bra{\alpha,0} \hat{\mathbf{I}}_{AS,out} \ket{\alpha,0}\\
							&= \cos^2(\Delta \phi) \vert \alpha \vert^2
		\end{aligned}
		\end{equation}
		
		which matches the classical description of a Michelson output.
		
		Following the same methods to calculate the radiation pressure, photon number variance is 
		\begin{equation} 
		\begin{aligned}
		\Delta \mathbf{I} 	&= \sqrt{\langle \mathbf{\hat{I}}^2 \rangle  - \langle \mathbf{\hat{I}} \rangle^2)} \\
							&= \vert \alpha \vert \big[ \cos^2(\Delta \phi)\ \big] 
		\end{aligned}
		\end{equation}
		
		\begin{equation}
		I = N \cos^2(\Delta \phi)
		\end{equation}
		
		\begin{equation}
		\frac{\partial I}{\partial \phi} = - 2 \vert \alpha \vert^2\cos(\Delta \phi) \sin(\Delta \phi)
		\end{equation}
		
		\begin{equation}
		\delta \phi = \sqrt{\frac{\hbar \omega}{ P_{in}}} \cot(\Delta \phi)
		\end{equation}
		Here $\delta \phi$ is the microscopic change in phase due to shot noise, whereas $\Delta\phi$ is the DC offset in the arm lengths to begin with. So in general, the shot noise contribution is dependent on the amount of light present at the anti-symmetric port and this will vary depending on what type of read out that is implemented.
		
		\begin{equation}
		\tilde{h}_{SN}(f) = 8.2\times 10^{-22} \; \bigg[ \frac{\text{Strain}}{\sqrt{\text{Hz}}}\bigg]
		\end{equation}
		Unsurprisingly, the variance is proportional to the amount of photons present and the phase difference between the interferometer arms.
		Interestingly, there are a few subtleties associated with measuring the noise, it is assumed here that the interferometer readout is a simple photodetector located at the antisymmetric port.  However, is it possible to measure the light at two different phases (90 degrees apart) and subtract the results.   Also, in section \ref{michelson}, there is a freedom to choose the readout scheme of the interferometer (heterodyne or homodyne) which will also affect the overall quantum noise.

		Losses!?!?!?!?!?!?
		
		\subsubsection{Seismic Noise}
		Seismic noise will be the low frequency barrier for all terrestrial gravitational-wave detectors.  By using seismic isolation platforms and quadruple suspensions, the noise contribution can be attenuated for frequencies larger than ~1Hz.
		
		Compare low and high noise microseisms
		[Blair: Norna]
		
		\subsubsection{Thermal Noise}
		Suspension Thermal Noise [Saulson and Gaby]
		Substrate Thermal Noise [\cite{Saulson}]
		Coating Thermal Noise [G Harry]
		
		\subsubsection{Newtonian Noise}
		Although there are some noise sources that can be reduced using increasingly complicated techniques such as more levels of seismic isolation or quantum nondemolition devices as shown above. [Saulson and Hughes/Thorne]
		
		Newtonian noise that is caused by fluctuating gravitational fields arising from the wave motion of the ground is one that cannot easily be reduced but possibly be measured and fed-forward or subtracted.[Driggers]
		
	
	\section{Squeezed States of Light}


	Virtually all undergraduate quantum mechanics textbooks include a section on the harmonic oscillator[Shankar, Sakura].  An interesting result when solving the Schrodinger equation using creation ($\hat{a^{\dagger}}$) and annihilation $\hat{a}$ operators is the existence of a non-zero energy ground state. 
	
	This effect leads to Quantum Noise, which is a fundamental source that can be improved by modifying the quantum vacuum using correlated photons and injecting the states of light into the antisymmetric port of the interferometer.  Within the LIGO community, this procedure of modifying quantum vacuum is called squeezing. Caves analytically derived the effects of quantum noise on the interferometer as well as the improvement due to	
	
	Losses!?!?!?!?!?!?
	References: [Caves, Dwyer, Kwee, Miao]
	